---
title: "LoadPancreasImages"
author: "Nicolas Damond"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "./IMC")
```

```{r load-packages, message=FALSE}
library(SingleCellMapper)
```

# Download the images and load them as an ImageList object
Here, a subset of 100 images from the pancreas IMC dataset is downloaded.

! The updated Mendeley dataset is not published yet, can be temporarily accessed from https://data.mendeley.com/datasets/cydmwsfztj/draft?a=e02ab17c-c014-4338-b379-101424e1c64d 
```{r load-images}
# Download the zipped folder image and unzip it
# url.images <- ("...")
# download.file(url.images, destfile = "images.zip")
# unzip("images.zip")
# remove.file("images.zip")

# (Temporary code to unzip the images, will be removed when the Mendeley dataset is public)
fn.images <- ("/home/nicolasd/Data/CM2017/mendeley/ImageSubset.zip")
unzip(fn.images)

# Load the images as an ImageList object
images <- loadImages("./", pattern='_full_clean.tiff')
images
```

# Download the cell masks and load them as an ImageList object
```{r load-masks}
# Download the zipped folder image and unzip it
# url.masks <- ("...")
# download.file(url.masks, destfile = "masks.zip")
# unzip("masks.zip")
# remove.file("masks.zip")

# (Temporary code to unzip the masks, will be removed when the Mendeley dataset is public)
fn.masks <- ("/home/nicolasd/Data/CM2017/mendeley/Masks.zip")
unzip(fn.masks)

# Load the images as an ImageList object
masks <- loadImages("./", pattern='_full_mask.tiff')
masks
```

# Load panel data
The panel contains antibody-related metadata.
The channel-mass file is used to match panel information and image stack slices.
```{r load-panel}
# Import panel
# url.panel <- ("...")
# download.file(url.panel, destfile = "panel.csv")
# panel <- read.csv(panel.csv)

# Import channel-mass file
# url.panel <- ("...")
# download.file(url.panel, destfile = "panel.csv")
# panel <- read.csv(panel.csv)

# Temporary code
panel <- read.csv("/home/nicolasd/Data/CM2017/mendeley/Panel.csv")
channel.mass <- read.csv("/home/nicolasd/Data/CM2017/mendeley/ChannelMass.csv", header=F)
```

## Scale masks
The masks are 16-bit images and need to be re-scaled.
```{r scale-masks}
masks <- endoapply(masks, function(x) x*(2^16-1))
```

## Add image names for images and masks
This information is stored in the metadata columns of the ImageList objects
and is used by SingleCellMapper to match single cell data, images and mask
```{r add-image-names}
mcols(images)$ImageName <- gsub("_a0_full_clean", "", names(images))
mcols(masks)$ImageName <- gsub("_a0_full_mask", "", names(masks))
```

# Add protein short names as channel names
```{r add-channel-names}
# Match panel and stack slice information
panel <- panel[panel$full == 1,]
panel <- panel[match(channel.mass[,1], panel$MetalTag),]

# Add channel names to the  image stacks ImageList object
channelNames(images) <- panel$shortname
```

# Save the ImageList objects
```{r save}
# save(images)
# save(masks)
```

