---
title: "Import of toy data"
author: "Nicolas Damond"
editor_options:
  chunk_output_type: inline
---

# Goal

Read in data exported from CellProfiler and store it as a SingleCellExperiment
For more information, see https://github.com/BodenmillerGroup/IMCWorkflow 

In this script, we will create a minimal toy dataset primarily used in the examples, unit tests and vignette of the \link{SingleCellMapper} package.

```{r load-packages, message=FALSE}
library(SingleCellExperiment)
library(SingleCellMapper)
```

# Load the datasets

## Load CellProfiler output and panel file

- The Cells.csv and Image.csv files are measurements exported from CellProfiler
- The panel.csv has to be modified manually for each experiment according to the antibody panel that was used.

```{r read-in-data}
cells <- read.csv(file = "~/switchdrive/Institution/SingleCellMapper/Data/raw_data/cpoutput/Cells.csv", stringsAsFactors = FALSE)
image <- read.csv(file = "~/switchdrive/Institution/SingleCellMapper/Data/raw_data/cpoutput/Image.csv", stringsAsFactors = FALSE)
panel <- read.csv(file = "~/switchdrive/Institution/SingleCellMapper/Data/raw_data/panel.csv", stringsAsFactors = FALSE)
```

# Load cell-specific, image-specific and feature metadata

## List of cell features measured by CellProfiler

```{r cell-specific-metadata}
unique(sub("_c[0-9]*$", "", colnames(cells)))
```

## Load relevant cell-specific metadata

```{r cell-metadata}
cell.metadata <- DataFrame(ImageNb = cells$ImageNumber,
                           CellNb = cells$ObjectNumber,
                           Pos_X = cells$Location_Center_X,
                           Pos_Y = cells$Location_Center_Y,
                           Area = cells$AreaShape_Area,
                           MajorAxisLength = cells$AreaShape_MajorAxisLength,
                           MinorAxisLength = cells$AreaShape_MinorAxisLength)
```

## List of image features measured by CellProfiler

```{r image-specific-metadata}
colnames(image)
```

## Load relevant image-specific metadata

```{r image-metadata}
image.metadata <- DataFrame(ImageNb = image$ImageNumber,
                            ImageName = image$FileName_ImageStack,
                            MaskName = image$FileName_Mask)
```

# Merge cell and image metadata

```{r}
cell.metadata <- merge(cell.metadata, image.metadata, by="ImageNb")
```

# Add rownames

```{r cell-rownames}
rownames(cell.metadata) <- paste(sub(".tiff", "", cell.metadata$ImageName),
                                 cell.metadata$CellNb,
                                 sep = ".")
```

# Feature-specific metadata

Make sure that the frame number is the same in the image stack and in the dataset

```{r reorder-panel}
panel <- DataFrame(panel)
channel.mass <- read.csv("~/switchdrive/Institution/SingleCellMapper/Data/raw_data/images/A02_imc.csv", header = FALSE)
panel <- panel[match(channel.mass[,1], panel$MetalTag), ]
```

Add clean target names as rownames in the panel

```{r panel-rownames}
rownames(panel) <- panel$clean_Target
```

# Load single cell measurements

## Single cell intensity measurements

Here, we select the mean intensity

```{r select-counts}
counts <- cells[, grepl("Intensity_MeanIntensity_ImageStack", colnames(cells))]
```

## Scale counts

Count are sometimes scaled in CellProfiler

```{r scale-counts}
scale.factor <- image$Scaling_ImageStack[1]
counts <- counts * scale.factor
```

## Reorder the counts channels (based on channel number)

```{r reorder-counts}
channelNumber <- as.numeric(sub("^.*_c", "", colnames(counts)))
counts <- counts[, order(channelNumber, decreasing = FALSE)] 
```

# Create the SingleCellExperiment object

## Create SCE

```{r create-SCE}
sce <- SingleCellExperiment(assays = list(counts = t(counts)))
```

## Add transformed counts

exprs = asinh-transformed

```{r}
assay(sce, "exprs") <- asinh(counts(sce)/1)
```

## Set rownames

```{r create-SCE}
rownames(sce) <- rownames(panel)
colnames(sce) <- rownames(cell.metadata)
```

### Store metadata
```{r store-metadata}
colData(sce) <- cell.metadata
rowData(sce) <- panel
sce
```

### Save SCE

```{r save-SCE}
saveRDS(sce, "~/switchdrive/Institution/SingleCellMapper/Data/raw_data/sce.rds")
```



