---
title: "Figure 4"
author: "Nils Eling and Nicolas Damond"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script reproduces the analysis performed in Figure 4.
Here, we will load the libraries and data for this figure:

```{r load-libraries-and-data, message=FALSE}
library(cytomapper)
library(dplyr)

sce <- readRDS("data/PancreasData/pancreas_sce.rds")
images <- readRDS("data/PancreasData/pancreas_images.rds")
```

Similar to Figure 3, we will now order the images based on the mean intensity of all PIN positive pixels. 

```{r order-by-beta-cells}
PIN_mean <- unlist(lapply(images,
                   function(x){
                     cur_x <- x[,,"PIN"]
                     mean(cur_x[cur_x > 0])
                   }))
```

Here, we will visualize the marker pro-insulin (PIN) across all images.
Due to the progressive decline in beta cell function, we expect PIN expression to be reduced in Long-Duration patients.
However, a similar analysis can be performed to visually detect image-to-image irregularities in background stain.

We will normalize across all images to keep the differences in staining intensity between images intact.

```{r plotCellTypes, fig.width=120, fig.height=120}
# Specify the image order
cur_order <- order(PIN_mean, decreasing = TRUE)

# Collect metadata for images
cur_meta <- as_tibble(colData(sce)) %>%
  group_by(ImageNumber) %>%
  summarise_at(vars("ImageName", "ImageFullName", "stage"), unique) %>%
  mutate(ImageFullName = gsub(".tiff", "", ImageFullName)) %>%
  as.data.frame()

rownames(cur_meta) <- cur_meta$ImageFullName
cur_meta <- cur_meta[names(PIN_mean)[cur_order],]

images <- normalize(images)
images <- normalize(images, inputRange = c(0, 0.2))

plotPixels(image = images[cur_order],
          colour_by = "PIN",
          scale_bar = list(length = 100,
                           label = "",
                           colour = "white"),
          colour = list(PIN = c("black", "yellow")),
          legend = list(margin = 100,
                        colour_by.title.cex = 6,
                        colour_by.labels.cex = 4),
          image_title = list(text = cur_meta$stage,
                             colour = "white",
                             cex = 3))

# Save plot
plotPixels(image = images[cur_order],
          colour_by = "PIN",
          scale_bar = list(length = 100,
                           label = "",
                           colour = "white"),
          colour = list(PIN = c("black", "yellow")),
          legend = list(margin = 100,
                        colour_by.title.cex = 8,
                        colour_by.labels.cex = 4),
          image_title = list(text = cur_meta$stage,
                             colour = "white",
                             cex = 3),
          save_plot = list(filename = "docs/final_figures/main/Fig_4.png"))
```

Here, we see the progressive decline in beta cell fractions. 
The images are ordered based on T1D stage: Non-diabetic, Onset and Long-Duration, as expected.
We further see an increased morphological irregularity in Long-Duration islets.

