---
title: "Running cytomapper shiny and Supplementary Figure 4"
author: "Nils Eling and Tobias Hoch"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script offers an example on how to run the `cytomapperShiny` function to gate and visualize cells on images.
For more information, please refer to `?cytomapperShiny` and the `Help` section within the Shiny app. 
We have taken screenshots from this shiny app to generate Supplementary Figure 4.

Here, we will read in the previously generated data objects and directly pass them into the `cytomapperShiny` function.

```{r load-libraries-and-data, message=FALSE}
library(cytomapper)

sce <- readRDS("data/PancreasData/pancreas_sce.rds")
masks <- readRDS("data/PancreasData/pancreas_masks.rds")
images <- readRDS("data/PancreasData/pancreas_images.rds")
```

## Classic hierarchical gating

When not supplying a segmentation mask or multi-channel image object, the `cytomapperShiny` function can be used to perform classic hierarchical gating similar to `FlowJo` on cell-specific expression values:

```{r cytomapperShiny-1}
if (interactive()) {
  cytomapperShiny(object = sce, img_id = "ImageName", cell_id = "CellNumber")
}
```

## Visualizing segmentation masks

Next, we will only supply the `SingleCellExperiment` object and the `CytoImageList` segmentation mask object to the Shiny function.
In this setting, segmentation masks are coloured (i) based on selected marker expression and (ii) based on the gated cells.

```{r cytomapperShiny-2}
if (interactive()) {
  cytomapperShiny(object = sce, mask = masks, 
                  cell_id = "CellNumber", img_id = "ImageName")
}
```

## Visualizing multi-channel images

In the third setting, we will gate cells based on their mean expression counts but outline the selected cells on composite images.
For this, we will need to supply the `SingleCellExperiment` object, a `CytoImageList` object containing segmentation masks and a `CytoImageList` object containing multi-channel images.

```{r cytomapperShiny-3}
if (interactive()) {
  cytomapperShiny(object = sce, mask = masks, image = images, 
                  cell_id = "CellNumber", img_id = "ImageName")
}
```



